---
title: "↯ Estimating Regression Discontinuity Designs (RDD) "
subtitle: "A `rough` replication using simulated data following the analyses in Neal, 2024"
author: "EDS 241"
format:
  html:
    theme: sketchy
date: "2025-02-04"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

------------------------------------------------------------------------

### RDD regression equation:

$Y_i = \beta_0 + \beta_1 D_i + \beta_2 X_i + \beta_3 (D_i \cdot X_i) + \epsilon_i$

-   $Y_i$: Outcome variable (`Deforestation rate`)
-   $D_i$: Treatment indicator (`Protected area`)
-   $X_i$: Running variable (`Distance to Protected Area Border (km)`) Has to be continuous
-   $D_i \cdot X_i$: Interaction term that allows slope vary across threshold
-   $\epsilon_i$: Error term

------------------------------------------------------------------------

### Load packages

------------------------------------------------------------------------

```{r}
library(tidyverse)
library(rdrobust)
library(here)
library(jtools)
library(janitor)
library(patchwork)
```

------------------------------------------------------------------------

### Read in simulated data to `roughly` replicate analyses in Neal, 24

```{r}

sim_data <- read_csv(here("data", "Simulated_Deforestation_Data5.csv")) 

```

------------------------------------------------------------------------

### Visualize the discontinuity using binned means (`bin size = 1`)

```{r}
data_binned <- sim_data %>%
  mutate(distance_bin = cut(distance, breaks = seq(-20, 20, by = 1), include.lowest = TRUE)) %>%
  group_by(distance_bin, protected) %>%
  summarize(
    avg_distance = mean(distance), # Averaged binned distance
    avg_deforest = mean(deforest),
    .groups = "drop"
  )
```

------------------------------------------------------------------------

### Plot using binned data

```{r}

ggplot(data_binned, aes(x = avg_distance, y = avg_deforest, color = as.factor(protected))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = protected), se = TRUE) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "Regression Discontinuity Plot (Binned by Distance)",
       x = "Running Variable (Distance)", y = "Deforestation Rate",
       color = "Protected Area") +
  theme_minimal()
```

------------------------------------------------------------------------

### Run RDD analysis using OLS

```{r}

rdd_ols <- lm(
    deforest ~  # outcome
    protected + # treatment effect 
    distance +  # running variable
    protected*distance + # allows slope to vary
    slope_cat + road_cat + water_cat + soil_cat, # <<<CONTROLS (simplify the analysis)
              data = sim_data)

# Display summary of regression results
summ(rdd_ols, digits = 3, 
     model.info = FALSE, model.fit = FALSE)
```

::: callout-tip
🚫 P-values (`p`) & standard errors (`S.E.`) are never zero!

\*Output values are printed `0.000` due to rounding settings
:::

::: callout
`NOTE`: We have modern estimation methods that do a better job at estimating a *robust* RDD treatment effect. However, an OLS analysis is often included for reference/comparison as a robustness check.
:::

------------------------------------------------------------------------

### Estimate & Visialize RDD using {`rdrobust`}

📜 [Documentation - rdrobust](https://cran.r-project.org/web/packages/rdrobust/rdrobust.pdf)

------------------------------------------------------------------------

**RDD Robust Estimation Method (`local polynomial regression`)**:

Local polynomial regression is a method used to give more weight to observations near a specific point— in this case, the RDD threshold. Instead of using OLS, it fits separate non-linear regressions on either side of the cutoff using a subset of the data near the cutoff (i.e., `bandwidth`).

**Interpreting output**:

*Default estimation options used by the `rdrobust()` function*:

-   Bandwidth Optimization (`BW type: mserd`): Bandwidth is optimized to balance accuracy & bias.
-   Bandwidth Estimate (`BW est. (h) = 5.729`): The estimated range around the cutoff used to subset the data to estimate the treatment effect.
-   Kernel (`Triangular`): Gives higher weight to data points close to the cutoff.
-   Variance Estimation (`VCE method: NN`): Instead of assuming equal variance across all observations, the error estimates are adjusted to account for variability near the cutoff.

------------------------------------------------------------------------

**Take a random sample (To adjust for memory-limit & speed)**

```{r}
global_samp <- sim_data %>%
     sample_n(size = nrow(sim_data) * 1.0) # <<< e.g., .5 for 50%
```

------------------------------------------------------------------------

### Estimate Global RDD

```{r}

global_rdd <- rdrobust(
  y = global_samp$deforest,
  x = global_samp$distance,
  covs = global_samp %>% select(slope_cat, road_cat, water_cat, soil_cat),
  c = 0, # Cutoff at 0 (protected area boundary)
  #p = 1, # Local linear regression
  kernel = "triangular"
)

# Print summary of results
summary(global_rdd)
```

------------------------------------------------------------------------

**Visualize the RDD discontinuity using `rdplot()`**:

This plot presents the local polynomial regression curves fit on either side of the cutoff.

```{r}
rdplot(
  y = sim_data$deforest,
  x = sim_data$distance,
  c = 0,
  binselect = "es",
  title = "Regression Discontinuity - Global Deforestation Rate",
  x.label = "Distance to Protected Area Border (km)",
  y.label = "Deforestation Rate"
)
```

### Estimate separate RDD models by `country` using function

> Country levels: Brazil, DRC, Malaysia, Indonesia, Canada, Australia

::: callout-tip
What is an iterator or loop function?

`lapply()` loops across the input levels for `country` and applies the function `run_country_rdd`
:::

```{r}

run_country_rdd <- function(country_name) {
  df_country <- sim_data %>% filter(country == country_name)

    rdd_model <- rdrobust(
      y = df_country$deforest,
      x = df_country$distance,
      covs = df_country %>% select(slope_cat, road_cat, water_cat, soil_cat),
      c = 0, 
      p = 1,
      kernel = "triangular"
    )
}

# Apply the function to all countries
rdd_6country <- lapply(unique(sim_data$country), run_country_rdd)

# Print summary for Malaysia
summary(rdd_6country[[3]]) 
```

### Generate country-level discontinuity plots

```{r, results="hide", fig.show="hide"}

# Create list to store plots
plot_list <- list()

# Loop across 6 countries and plot
for (country_name in unique(sim_data$country)) {
  df_country <- sim_data %>% filter(country == country_name)
  
 p <- rdplot(
      y = df_country$deforest,
      x = df_country$distance,
      c = 0,
      binselect = "es",
      title = paste(country_name),
      x.label="", y.label = "")
  
 p <- p$rdplot +
     labs(x=" ",y="")
 
  plot_list[[country_name]] <- p  # Store each plot in the list
  }
```

```{r, results='hide'}
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 2) +  
  plot_annotation(
    title = "Deforestation Rate",
    caption = "Source: Simulation Data"
  )
```

### Print combined RDD plots

```{r}
final_plot + 
  labs(x = "Distance to Protected Area Border (km)")
```
